CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-08 codecheck-paper.tex
related work section
(I think this will need further tweaking)

We use freely available infrastructure, GitHub and Zenodo, to run our
system.  The \texttt{codecheckers} \textbf{GitHub}
organisation\footnote{ \url{https://github.com/codecheckers}} contains
projects for managing the project website, the codecheckers community
and its discussions, code repositories, and the main register of
CODECHECKs. Both the website\footnote{
  \url{https://codecheck.org.uk/}} and the register\footnote{
  \url{https://codecheck.org.uk/register}} are hosted as GitHub
pages. The register is a single table in CSV format that connects the
certificate identifier with the repository associated with a specific
conducted CODECHECK. Each of these repositories, which currently can
be hosted on GitHub or Open Science Framework (OSF), contains the
CODECHECK metadata file \texttt{codecheck.yml}. The register further
contains a column for the type of check, e.g.,~community, journal, or
conference, and the respective GitHub issue where communications and
assignments around a specific check are organised. No information is
duplicated between the register and the metadata files. The continuous
integration infrastructure of GitHub, GitHub Actions, is used to
automate the register for viewing (HTML) and for integration with
other services (CSV, JSON).  \textbf{Zenodo} is our preferred open repository for
storing certificates. It mints DOIs for deposits and ensures
long-term availability of all digital artefacts related to the
project. The CODECHECK community on Zenodo
\footnote{\url{https://zenodo.org/communities/codecheck/}} includes
certificates as well as the regularly archived register
\cite{codecheck_register_jan2021}.

@@

The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
established a ``Replicated Computational Results'' (RCR) review process
\cite{heroux_editorial_2015}, where ``replicable'' means the same thing as what we call ``reproducible''.
A search on Crossref\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2}; search executed
on 2020-12-10.} shows that 15 RCR Reports have been published and the 
% Number of found TOMAS RCR reports:
% 8: https://scholar.google.de/scholar?start=10&q=Replicated+Computations+Results+%22(RCR)+Report%22&hl=de&as_sdt=0,5
% 15: https://search.crossref.org/?q=Replicated+Computations+Results+%22%28RCR%29+Report%22&from_ui=yes&publication=ACM+Transactions+on+Mathematical+Software
% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
workflow is being
extended to the ACM journal \emph{Transactions on Modeling and Computer 
Simulation (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
The TOMS RCR follows the same principles as those used by CODECHECK (principles 1-4), which were independently developed as we only later learned about TOMS RCR.
Here, we provide an extensive discussion of the possible variants of CODECHECK,
while TOMS RCR realises one particular workflow for reproducibility 
reviews for one particular journal. The TOMS editorial
\cite{heroux_editorial_2015} shares similar concerns about
selection of reviewers, as we discussed above. Unlike existing CODECHECK certificates, 
the RCR reports, which are very similar to CODECHECK certificates, undergo editorial review, but this could eventually become
part of the CODECHECK processes at journals.
The publication of the RCR report is seen as a reward for the efforts of
the reproducing person, while the potential for this motive to be a conflict of interest is acknowledged.
TOMS also recognises reviewer activity in a 
partnership with Publons\footnote{
\url{https://authors.acm.org/author-services/publons}},
though it is unclear whether the author of an RCR report receives both the report's
authorship and the reviewer recognition. In any case, this is different
from CODECHECK, which specifically aims to give codecheckers recognition as reviewers.
In our view, the reviewer role removes the possible conflict of interest while
keeping the public acknowledgement.
Specific to the field of mathematics, the RCR is also expected to apply a
review of the software itself if the system it runs on cannot be evaluated by
an independent party.
The TOMS RCR creators concur with the importance of communication, expect
collaboration between author and RCR reviewers,
share the considerations around reviewer selection, and also put trust
in reviewer judgement over numerical bit-wise perfection.
A key difference is that for TOMS RCR, authors opt-in with an \emph{RCR Review Request} and
the RCR reports are published in the TOMS journal next to the actual papers.

A number of journals provide special article types for reproductions of papers
published in the journal.
\emph{Information Systems} has an invited 
reproducibility paper track\footnote{
\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}.}.
\emph{Nature Machine Intelligence} recently introduced a new type of article,
the reusability report \cite{noauthor_research_2020}.
Inspired by the detailed and nuanced submissions to a reproducibility 
challenge, the reusability report focuses on the exploration of robustness
and generalizability of the original paper's claims
\cite{noauthor_research_2020}. This answers the specific community's 
challenges around computational reproducibility and also values these kinds
of contributions as independent publications, which goes beyond the goals 
of CODECHECK.
The journal \emph{Cortex} has a special article type 
\emph{Verification Reports}, which are actually about replication of results
but are very well designed/reasoned \cite{chambers_verification_2020}.
In a similar vein, the CODECHECK certificates
could also be published as a special article type within journals.

@@

Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline and 
infrastructure for reproducible computational research with its own tools and
API. Their specific packaging format, provenance record, and cryptographic
\emph{Verifiable Result Identifier} would indeed provide excellent
reproduciblity. However, the system is also complex and has not been adopted by
any publisher as far as we know; also, the system is indeed not open source itself.
In comparison, CODECHECK is less powerful but also much more flexible and
less dependent on specific tools or infrastructure. If data and code are 
deposited properly, i.e.,~very unlikely to disappear, then the certificate's 
DOI is practically close to the cryptographic identifier.

@@

No approach besides CODECHECK spans across journals and is
intentionally designed to be picked up and implemented by many different 
journals or events and to build a community of codecheckers. CODECHECK shares
its interdisciplinary nature with other community initiatives concerned with 
reproducibility awareness, education, and support, such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaboration with the preprint server
\emph{arXiv} on providing data and code supplements for machine learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
% Papers with Code now also does more disciplines than just ML.
Likewise, different disciplines and journals provide reproducibility
checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or GIScience \cite{nust_agile_2019}, which naturally share some
aspects while addressing particularities as well as addressing researchers
from different fields. Regarding the education and guidance for authors, 
we see CODECHECK's role as referencing and linking educational efforts
and helpful material, not as creating and maintaining such content.

>>>>

We use freely available infrastructure, GitHub and Zenodo, to run our
system.  The \texttt{codecheckers} \textbf{GitHub}
organisation\footnote{ \url{https://github.com/codecheckers}} contains
projects for managing the project website, the codecheckers community
and its discussions, code repositories, and the main register of
CODECHECKs. Both the website\footnote{
  \url{https://codecheck.org.uk/}} and the register\footnote{
  \url{https://codecheck.org.uk/register}} are hosted as GitHub
pages. The register is a single table in CSV format that connects the
certificate identifier with the repository associated with a specific
conducted CODECHECK. Each of these repositories, which currently can
be hosted on GitHub or Open Science Framework (OSF), contains the
CODECHECK metadata file \texttt{codecheck.yml}. The register further
contains a column for the type of check, e.g.,~community, journal, or
conference, and the respective GitHub issue where communications and
assignments around a specific check are organised. No information is
duplicated between the register and the metadata files. The continuous
integration infrastructure of GitHub, GitHub Actions, is used to
automate the register for viewing (HTML) and for integration with
other services (CSV, JSON).  \textbf{Zenodo} is our preferred open repository for
storing certificates. It mints DOIs for deposits and ensures
long-term availability of all digital artefacts related to the
project. The CODECHECK community on Zenodo
\footnote{\url{https://zenodo.org/communities/codecheck/}} includes
certificates and the regularly archived register
\cite{codecheck_register_jan2021}.

@@

The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
recently established a ``Replicated Computational Results'' (RCR) review process
\cite{heroux_editorial_2015}, where ``replicable'' is the same as our
use of ``reproducible''.

Fifteen RCR Reports have been published so far
\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2};
  search executed on 2020-12-10.}
% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
and the process is being extended extended to the ACM journal
\emph{Transactions on Modeling and Computer Simulation
  (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
The TOMS RCR follows CODECHECK principles 1-4, although our work was
independently developed of theirs.  Here, we discuss variants of
CODECHECK, while TOMS RCR realises one particular workflow for
reproducibility reviews for one particular journal. The TOMS editorial
\cite{heroux_editorial_2015} shares similar concerns about selection
of reviewers, as we discussed above. Unlike existing CODECHECK
certificates, the RCR reports, which are similar to CODECHECK
certificates, undergo editorial review, but this could eventually
become part of the CODECHECK processes at journals.  The publication
of the RCR report is seen as a reward for the efforts of the
reproducing person, while the potential for this motive to be a
conflict of interest is acknowledged.  TOMS also recognises reviewer
activity in a partnership with Publons\footnote{
  \url{https://authors.acm.org/author-services/publons}}.
%, but it is
unclear if the author of an RCR report receives both the report's
%authorship and the reviewer recognition. 
CODECHECK however specifically aims to give codecheckers
recognition as reviewers.  In our view, the reviewer role removes the
possible conflict of interest while keeping the public
acknowledgement.  Specific to the field of mathematics, the RCR is
also expected to apply a review of the software itself if the system
it runs on cannot be evaluated by an independent party.  The TOMS RCR
creators concur with the importance of communication, expect
collaboration between author and RCR reviewers, share the
considerations around reviewer selection, and also put trust in
reviewer judgement over numerical bit-wise perfection.  A key
difference is that for TOMS RCR, authors opt-in with an \emph{RCR
  Review Request} and the RCR reports are published in the TOMS
journal next to the actual papers.
%% SJE: think this could be tightened a bit, and some of it I don't
%% fully understand.

Several journals provide special article types for reproductions of
published papers.  \emph{Information Systems} has an
invited reproducibility paper track\footnote{
  \url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}.}.
\emph{Nature Machine Intelligence} recently introduced a new type of
article, the reusability report \cite{noauthor_research_2020}.
Inspired by the detailed and nuanced submissions to a reproducibility
challenge, the reusability report focuses on the exploration of
robustness and generalizability of the original paper's claims
\cite{noauthor_research_2020}. This answers the specific community's
challenges around computational reproducibility and also values these
kinds of contributions as independent publications, which goes beyond
the goals of CODECHECK.  The journal \emph{Cortex} has a special
article type \emph{Verification Reports}, which are actually about
replication of results and are very well designed/reasoned
\cite{chambers_verification_2020}.  In a similar vein, the CODECHECK
certificates could also be published as a special article type within
journals.

@@

Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline
and infrastructure for reproducible computational research with its
own tools and API. Their specific packaging format, provenance record,
and cryptographic \emph{Verifiable Result Identifier} would indeed
provide excellent reproduciblity. However, the system is also complex
and since its creation in 2011 we are not aware of any publisher using
it; also, the system is not open source itself.  In comparison,
CODECHECK is less powerful but also much more flexible and less
dependent on specific tools or infrastructure. If data and code are
deposited properly, i.e.,~very unlikely to disappear, then the
certificate's DOI is practically close to the cryptographic
identifier.

@@

No approach besides CODECHECK spans across journals and is
intentionally designed to be adopted across journals or events and to
build a community of codecheckers. CODECHECK shares its
interdisciplinary nature with other community initiatives concerned
with reproducibility awareness, education, and support, such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or
Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaboration with the preprint server
\emph{arXiv} on providing data and code supplements for machine
learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
and runs a reproduciblity
challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
% Papers with Code now also does more disciplines than just ML.
Likewise, different disciplines and journals provide reproducibility
checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or
GIScience \cite{nust_agile_2019}, which naturally share some aspects
while addressing particularities as well as addressing researchers
from different fields. Regarding the education and guidance for
authors, we see CODECHECK's role as referencing and linking
educational efforts and helpful material, not as creating and
maintaining such content.
